A User-Extensible and Adaptable Parser Architecture

### 2008-08-20 doc/sgai-2008.tex
Final suggestions from Carl. Remaining corrections from my proofreading. Resolved all XXX items. Improved bibliography entries.

% $Id$
% XXX REMOVE draft OPTION
\documentclass[draft]{svmult}

@@

% XXX REMOVE THESE AND CHECK RESULTING LENGTH
%\setlength{\floatsep}{0pt}
%\setlength{\intextsep}{0pt}
%\setlength{\textfloatsep}{0pt}

@@

\title*{A User-Extensible and Adaptable Parser Architecture}
\author{John Tobin and Carl Vogel}
\institute{John Tobin \at{} School of Computer Science and Statistics,
Trinity College, Dublin 2, Ireland. \newline{} \email{tobinjt@cs.tcd.ie}
\and{} Carl Vogel \at{} School of Computer Science and Statistics,
Trinity College, Dublin 2, Ireland. \newline{} \email{vogel@cs.tcd.ie}}
\maketitle

@@

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture: instead of writing one monolithic
function which must be modified to support new behaviour, with all the
attendant risks of adversely affecting the existing parser, when a new
requirement arises an independent action can be written to satisfy it.  In
some cases the new action will require the cooperation of other actions,
e.g.\ to set or check a flag; there is a possibility of introducing failure
by modifying existing actions, but modifications will be smaller and occur
less frequently than with a monolithic architecture, thus failures should
be less likely, and will be easier to test for and diagnose.  The parser is
written in an object oriented style, allowing sub-classes to extend or
override actions in addition to adding new actions; because each action is
an independent procedure, the sub-class need only modify the action it is
overriding, rather than reproducing large chunks of functionality.

@@

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}; it is possible to convert between regexes and
\FA{}, so regex overlap could be detected by finding a non-empty
intersection of two \FA{}\@.  One formula for \FA{} intersection is given
in~\cite{intersection-of-NFA-using-Z}: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{}, so overlap detection may
require calculating the intersection of two \PDA{}\@.  \CFL{} are
equivalent in computational power to \PDA{}, and intersection of two
\CFL{s} is not closed, i.e.\ intersection can produce a language which
cannot be parsed by a \CFL{}, e.g.\ $a^{*}b^{n}c^{n} \cap a^{n}b^{n}c^{*}
\rightarrow a^{n}b^{n}c^{n}$, so the results of \PDA{} intersection may not
be meaningful in some cases.  Detecting overlap between $n$ regexes
requires $n(n-1)/2$ intersections to be calculated, resulting in $O(n^2x)$
complexity, where $O(x)$ is the complexity of calculating intersection.
This is certainly not a task to be performed on every parsing run:
detecting overlap amongst the Postfix log parser's \numberOFrules{} rules
would require calculating \numberOFruleINTERSECTIONS{} intersections.

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse very large log files.  The mail server which
generated the log files used in testing the Postfix log parser accepts
approximately 10,000 mails for 700 users a day.  Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  When generating the timing data used in this
section, \numberOFlogFILES{} log files (totaling 10.08 GB,
\numberOFlogLINEShuman{} log lines) were each parsed 10 times and the
parsing times averaged.  Saving results to the database was disabled for
the test runs, as the tests are aimed at measuring the speed of the Postfix
log parser rather than the speed of the database and the disks the database
is stored on.  The computer used for test runs was a Dell Optiplex 745
described in \tableref{Computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and was not
used for any other purpose while test runs were ongoing --- any services
not necessary for running the tests were disabled.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.

% XXX MAYBE CHANGE POSITIONING
\begin{table}[htbp]
    \caption{Computer used to generate statistics}
    \empty{}\label{Computer used to generate statistics}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Component  & Component used                                 \\
        \tablemiddleline{}%
        CPU        & 1 dual core 2.40GHz Intel\textregistered{}
                     Core\texttrademark{}2 CPU,                      
                     with 32KB L1 and 4MB L2 cache.                 \\
        RAM        & 2GB 667 MHz DDR RAM\@.                         \\
        Hard disk  & 1 Seagate Barracuda 7200 RPM 250GB SATA disk.  \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

At the time of writing the Postfix log parser has \numberOFrules{} different rules, with the
top 10\% matching the vast majority of log lines, and the remaining log
lines split across the other 90\% of the rules in a Power Law distribution
(as shown in \graphref{rule hits graph}).  Assuming that the distribution
of log lines is reasonably consistent over time, parser efficiency should
benefit from trying more frequently matching rules before those which match
less frequently.  To test this hypothesis three full test runs were
performed with different rule orderings:

>>>>

% $Id$
\documentclass{svmult}
%\documentclass[draft]{svmult}

@@

@@

\title*{A User-Extensible and Adaptable Parser Architecture}
\author{John Tobin and Carl Vogel}
\institute{John Tobin \email{tobinjt@cs.tcd.ie}
            \and{} Carl Vogel \email{vogel@cs.tcd.ie}
            \at{} School of Computer Science and Statistics,
                    Trinity College, Dublin 2, Ireland.
                    \newline{}
                    Supported by Science Foundation Ireland RFP
                    05/RF/CMS002.
            }

\maketitle

@@

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture: instead of writing one monolithic
function which must be modified to support new behaviour, with all the
attendant risks of adversely affecting the existing parser, when a new
requirement arises an independent action can be written to satisfy it.  In
some cases the new action will require the cooperation of other actions,
e.g.\ to set or check a flag.  There is a possibility of introducing
failure when modifying existing actions, but the modifications will be
smaller and occur less frequently than with a monolithic architecture, thus
failures will be less likely and will be easier to test for and diagnose.
The parser is written in an object oriented style, allowing sub-classes to
extend or override actions in addition to adding new actions; because each
action is an independent procedure, the sub-class need only modify the
action it is overriding, rather than reproducing large chunks of
functionality.

@@

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}; it is possible to convert between regexes and
\FA{}, so regex overlap could be detected by finding a non-empty
intersection of two \FA{}\@.  One formula for \FA{} intersection is given
in for example~\cite{intersection-of-NFA-using-Z}: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{}, so overlap detection may
require calculating the intersection of two \PDA{}\@.  \CFL{} are
equivalent in computational power to \PDA{}, and intersection of two
\CFL{s} is not closed, i.e.\ intersection can produce a language which
cannot be parsed by a \CFL{}, e.g.\ $a^{*}b^{n}c^{n} \cap a^{n}b^{n}c^{*}
\rightarrow a^{n}b^{n}c^{n}$, so the results of \PDA{} intersection may be
intractable in some cases.  Detecting overlap between $n$ regexes requires
$n(n-1)/2$ intersections to be calculated, resulting in $O(n^2x)$
complexity, where $O(x)$ is the complexity of calculating intersection.
This is certainly not a task to be performed on every parsing run:
detecting overlap amongst the Postfix log parser's \numberOFrules{} rules
would require calculating \numberOFruleINTERSECTIONS{} intersections.

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse very large log files.  The mail server which
generated the log files used in testing the Postfix log parser accepts
approximately 10,000 mails for 700 users a day.  Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  When generating the timing data used in this
section, \numberOFlogFILES{} log files (totaling 10.08 GB,
\numberOFlogLINEShuman{} log lines) were each parsed 10 times and the
parsing times averaged.  Saving results to the database was disabled for
the test runs, as the tests are aimed at measuring the speed of the Postfix
log parser rather than the speed of the database.  The computer used for
test runs was a Dell Optiplex 745 described in \tableref{Computer used to
generate statistics}; it was dedicated to the task of gathering statistics
from test runs, and was not used for any other purpose while test runs were
ongoing --- any services not necessary for running the tests were disabled.
Parsing all \numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.

\begin{table}[htbp]
    \caption{Computer used to generate statistics}
    \empty{}\label{Computer used to generate statistics}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Component  & Component used                                 \\
        \tablemiddleline{}%
        CPU        & 1 dual core 2.40GHz Intel\textregistered{}
                     Core\texttrademark{}2 CPU,                      
                     with 32KB L1 and 4MB L2 cache.                 \\
        RAM        & 2GB 667 MHz DDR RAM\@.                         \\
        Hard disk  & 1 Seagate Barracuda 7200 RPM 250GB SATA disk.  \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

At the time of writing the Postfix log parser has \numberOFrules{}
different rules: the top 10\% match the vast majority of log lines, with
the remaining log lines split across the other 90\% of the rules.  The
distribution resembles a Power Law distribution, as shown in \graphref{rule
hits graph}.  Assuming that the distribution of log lines is reasonably
consistent over time, parser efficiency should benefit from trying more
frequently matching rules before those which match less frequently.  To
test this hypothesis three full test runs were performed with different
rule orderings:
